---
title: "8_29_23 Coding Notes"
author: "Dan Schumacher"
format: html
editor: source
---

We begin by installing `tidyverse`.

```{r}
#|  eval: false
# ^ This makes it so you don't re install tidyverse everytime
install.packages('tidyverse')
```

Load tidyverse

```{r}
library(tidyverse)
```

## dataframes

R is strong bc dataframes are native to r (not like how pandas isn't native to python, new package called polars)

`mpg` is a dataframe from 'ggplot2' packages

This doesn't look that good

```{r}
print(mpg)
```

```{r}
head(mpg)
```

What is a data frame? A collection of items (in r they called vectors, these are collums) they must have the same length! the objects have to be same type <int>, <chr> etc. We use NA (not Null or .)

Useful metadata functions

```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
colnames(mpg)
class(mpg)
```

TBL vs. DFs

```{r}
mpg_df = as.data.frame(mpg)
class(mpg_df)
```

tbl by default only prints 10 rows. also tbl shows data types.

## Subsetting a dataframe

`select()` for columns

-   outputs the same type of data as input
-   number rows stays the same
-   plays nice with other functions
    -   `Starts_with()`
    -   `ends_with()`
    -   `contains()`

```{r}
select(.data = mpg, year)
```

```{r}
select(mpg, starts_with('c'))
```

```{r}
select(mpg, ends_with('y'))
```

```{r}
select(mpg, contains('e'))
```

### filtering rows

`filter()` for rows

```{r}
filter(.data = mpg, year == 2008)
```

can use multiple conditions as well the ***','*** means AND so better to write & explicitly.

```{r}
filter(mpg, year == 2008 , manufacturer == 'audi')
```

`slice()` lets you filter rows by index (inclusive)

```{r}
slice(.data = mpg, 10:15)
```
