---
title: "8_29_23 class notes"
author: "Dan Schumacher"
format: html
editor: source
---

This \^ is yaml (yet another markdown language)

change format to docx, or pdf if you want a different output after render.

General Notes ---------------------------------------------------------------------------------------------------------------------

if a doc can be opened without unzip, do that it saves memory don't open CSV files in Excel. It will mess up the file! especially the date file can get messed up

IDE is a shell that provides additional functionality (integrated developer environment)

Console is where you interact with R Terminal interacts with your machine \
Bottom right window has lots of tabs

file explorer

Help tab packages plots presentation

Top right has lots of tabs

environment (tells you what is active in session)

git history of commands run posit IDE (R studio) we will be using quarto documents

what is my working directory?

```{r}
getwd()
```

you can type ctrl-alt-i to make a new code block

`#| echo:` false makes hides the *code* that you write but keeps the *input*.

```{r}
#| echo: false
getwd() 
```

```{r}
```
